### 2019-10-29 Paper_shortened_doc.Rmd
DP submission

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. The present study investigates two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Inductive reasoning is thought to promote the internalization of parental expectations, including display rules for emotions, and might thus play a role in emotional adjustment development as well [@grusecImpactParentalDiscipline1994].

@@

Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010]. Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers. For various reasons, mothers and fathers sometimes appear to fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011]. In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage, and respond more supportively to, emotional expression, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for adolescents' emotional adjustment, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is some evidence that emotional adjustment is more negatively impacted by negative parenting from mothers, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship. Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.

@@

A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001]. Examining the childcare distribution instead, and controlling for total care, mitigates this confound. Some prior research has attempted to measure the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016]. When quantitative involvement has been measured in terms of absolute involvement, however, a "latent mean and difference" approach can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].

@@

table_cors <- rbind(table_between_cor, table_within_cor)
write.csv(table_cors, "table_cors.csv", row.names = FALSE)

reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")

reg_table(rbind(tabparenteff, tabchildeff), caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
reg_table(rbind(tabinv, tabinvmod), caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

>>>>

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. We examined two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Inductive reasoning is thought to promote the internalization of parental expectations, including display rules for emotions, and might thus play a role in emotional adjustment development as well [@grusecImpactParentalDiscipline1994].

@@

Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010]. Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers. For various reasons, mothers and fathers sometimes appear to fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011]. In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage, and respond more supportively to, emotional expression, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for adolescents' emotional adjustment, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship. We might therefore expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.

@@

A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001]. Examining the childcare distribution instead, and controlling for total care, mitigates this confound. Some prior research has attempted to measure the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016]. When quantitative involvement has been measured in terms of absolute involvement, latent means and differences can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].

@@

table_cors <- rbind(table_between_cor, table_within_cor)
write.csv(table_cors, "table_cors.csv", row.names = FALSE)
table_cors[is.na(table_cors)] <- ""
rownames(table_cors) <- NULL
#reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
tab <- rbind(tabparenteff, tabchildeff)
tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

### 2019-10-29 Paper_shortened_pdf.Rmd
DP submission

keywords          : "involvement; emotional adjustment; fathers; parenting; longitudinal"
wordcount         : "7361"

@@

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

@@

We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.

@@

table_cors <- rbind(table_between_cor, table_within_cor)
write.csv(table_cors, "table_cors.csv", row.names = FALSE)
table_cors[is.na(table_cors)] <- ""
rownames(table_cors) <- NULL
#reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
tab <- rbind(tabparenteff, tabchildeff)
tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers. This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a]. With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting. Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.

@@

In line with prior research, we found that parents raise boys and girls differently. Mothers and fathers both displayed more anger and inductive reasoning with sons than with daughters, and fathers displayed less warmth towards sons. These findings suggest that parents are stricter towards their sons. Parenting effects on children were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors. Child effects were moderated by sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male. When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.

@@

We also found substantial evidence that both parents showed more desirable patterns of qualitative parenting behaviors - behaviors that are positively associated with children's adjustment - when total joint involvement was higher. When mothers were more involved than fathers, however, mothers' expressed more warmth, whereas fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.

>>>>

keywords          : "involvement; emotional adjustment; fathers; parenting; longitudinal"
wordcount         : "7356"

@@

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : yes

@@

We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. Each wave was reviewed and approved by the
`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
Ethics Committee registered with the `r masked("National Health and Medical Research Council")`. Documentation of the ethics approval is available online at `r masked("https://growingupinaustralia.gov.au/sites/default/files/certification_of_ethics_approval_for_lsac.pdf")`. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.

@@

table_cors <- rbind(table_between_cor, table_within_cor)
write.csv(table_cors, "table_cors.csv", row.names = FALSE)
#table_cors[is.na(table_cors)] <- ""
rownames(table_cors) <- NULL
#reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
tab <- rbind(tabparenteff, tabchildeff)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
#tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

With regard to parenting effects, the only difference between mothers and fathers was for the predictive effect of warmth on emotional adjustment, which was evident only for mothers. This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a]. With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting. Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.

@@

In line with prior research, we found that parents raise boys and girls differently. Mothers and fathers both displayed more anger and inductive reasoning with sons than with daughters, and fathers displayed less warmth towards sons. These findings suggest that parents are stricter towards their sons. Parenting effects were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors. Child effects were moderated by sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male. When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.

@@

We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when total joint involvement was higher. Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.
